### 2024-04-09 03_notes/mathematical_notes.tex
First Implementation of TGR Estimate
- Basic Mathematical Derivation
- First implementation of Triple Gamma Regularization Estimate

\documentclass[12pt,a4paper]{article}
\usepackage[
	left 	= 2.54cm,
	right 	= 2.54cm, 
	top 		= 2.54cm,
	bottom 	= 2.54cm,
]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[OT1]{fontenc}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{caption}

\usepackage[style=apa, backend=biber]{biblatex}
\addbibresource[]{ref.bib}
\addbibresource[]{DevEcon_Project.txt}
\renewcommand{\baselinestretch}{1.25}

\usepackage[hidelinks]{hyperref}
\hypersetup{
	colorlinks = true,
	urlcolor   = blue,
	linkcolor  = black, 
	citecolor  = blue, 
}

\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{\slshape Unterweger}
\chead{}
\rhead{\slshape \nouppercase{\leftmark}}

\usepackage{titlesec,xcolor}
\titleformat{\section}{\bfseries}{\thesection}{0.5em}{}
\titlespacing{\section}{0pt}{3ex plus 1ex minus 0.2ex}{10pt}
\setlength{\headheight}{14.49998pt}

\usepackage{titlesec,xcolor}
\titleformat{\subsection}{\bfseries}{\thesubsection}{0.5em}{}
\titlespacing{\subsection}{0pt}{3ex plus 1ex minus 0.2ex}{10pt}
\setlength{\headheight}{14.49998pt}

\author{Lucas Paul Unterweger}
\title{Mathematical Notes MSc Thesis}
\begin{document}
\maketitle

\section{Important Prior Distributions}

\textbf{Normal Distribution}
\[
f(x|\mu, \sigma) = \frac{1}{\sigma\sqrt{2\pi}} e^{-\frac{1}{2}\left(\frac{x-\mu}{\sigma}\right)^2}
\]

\textbf{Gamma Distribution}
\begin{equation*}
f(x|\alpha,\beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} x^{\alpha-1} e^{-\beta x}, \quad x > 0
\end{equation*}
where $\alpha$ is the shape parameter, $\beta$ is the rate parameter, and $\Gamma(\cdot)$ is the gamma function.

\textbf{Beta Distribution}
\begin{equation*}
f(x|a,b) = \frac{x^{a-1}(1-x)^{b-1}}{B(a,b)}
\end{equation*}
where $x \in [0,1]$, $a > 0$, $b > 0$, and $B(a,b)$ is the Beta function defined as:
\begin{equation*}
B(a,b) = \frac{\Gamma(a)\Gamma(b)}{\Gamma(a+b)}
\end{equation*}
where $\Gamma$ is the gamma function.\\

\section{Other Functions}
\textbf{Confluent Hyper-Geometrics Function of the second kind (Tricomi's (confluent hypergeometric) function)}
\[
U(a,b,z) = \frac{1}{\Gamma(a)}\int_0^\infty e^{-zt}t^{a-1}(1+t)^{b-a-1}dt
\]
\textbf{Marginal Prior with $\phi^\xi = \frac{2c^\xi}{\kappa_B^2a^\xi}$}
\begin{align*}
p(\sqrt{\theta_j}|\phi^\xi, a^\xi, c^\xi) &= \frac{\Gamma(c^\xi + \frac{1}{2})}{\sqrt{2\pi \phi^\xi}\cdot B(a^\xi, c^\xi)}\cdot U\left(c^\xi + \frac{1}{2}, \frac{3}{2}-a^\xi, \frac{\theta_j}{2\phi^\xi}\right)\\
&\propto U\left(c^\xi + \frac{1}{2}, \frac{3}{2}-a^\xi, \frac{\theta_j}{2\phi^\xi}\right)
\end{align*}

\section{Bayesian Regression}
Consider a Bayesian regression model where the relationship between the independent variable $x$ and the dependent variable $y$ is modeled as:
\begin{equation*}
y = \beta_0 + \beta_1 x + \epsilon
\end{equation*}
where $\epsilon$ is the error term assumed to follow a normal distribution with mean $0$ and variance $\sigma^2$.

We assume the following conjugate prior distributions for the parameters:
\begin{align*}
\beta_0 & \sim \mathcal{N}(\mu_0, \tau_0^2) \\
\beta_1 & \sim \mathcal{N}(\mu_1, \tau_1^2) \\
\sigma^2 & \sim \text{Inv-Gamma}(\alpha, \beta)
\end{align*}

The likelihood function for the observed data $(x_i, y_i)$, $i = 1, \ldots, n$, is given by:
\begin{equation*}
L(\beta_0, \beta_1, \sigma^2 | x, y) = \prod_{i=1}^{n} f(y_i | \beta_0, \beta_1, x_i, \sigma^2)
\end{equation*}
where $f(y_i | \beta_0, \beta_1, x_i, \sigma^2)$ is the probability density function (PDF) of the normal distribution.

The posterior distribution of the parameters, denoted as $\pi(\beta_0, \beta_1, \sigma^2 | x, y)$, is proportional to the product of the likelihood function and the prior distributions:
\begin{equation*}
\pi(\beta_0, \beta_1, \sigma^2 | x, y) \propto L(\beta_0, \beta_1, \sigma^2 | x, y) \times \pi(\beta_0) \times \pi(\beta_1) \times \pi(\sigma^2)
\end{equation*}

The maximum a posteriori (MAP) estimation involves maximizing the log-posterior function:
\begin{equation*}
\ell(\beta_0, \beta_1, \sigma^2 | x, y) = \sum_{i=1}^{n} \log f(y_i | \beta_0, \beta_1, x_i, \sigma^2) + \log \pi(\beta_0) + \log \pi(\beta_1) + \log \pi(\sigma^2)
\end{equation*}
with respect to the parameters $\beta_0$, $\beta_1$, and $\sigma^2$.

The estimates of the parameters, denoted as $\hat{\beta}_0$, $\hat{\beta}_1$, and $\hat{\sigma}^2$, can be obtained by maximizing the log-posterior function.

\newpage

\section{Derviation Idea}
\begin{equation*}
post(\beta, \sigma^2 | x, y) \propto L(\beta, \sigma^2 | x, y) \times prior(\beta, \sigma^2)
\end{equation*}
\end{document}